\chapter{Acknowledgments}

\textit{Zurich, Switzerland - October 2024}

\vspace*{\baselineskip}
First and foremost I would like to thank Mutian Niu and Usman Arshad for hosting me in the Animal Nutrition Group at Strickhof and providing excellent guidance throughout the past months to enable steady progress in our interdisciplinary project despite unforeseen hurdles with the sparsity of our data. Moreover, I also express my gratitude to Boris Zandona, Meiqing Wang and Elli Broxham Stahl for accommodating all my side requests regarding logistics and infrastructure.

\vspace*{\baselineskip}
Furthermore, this work could not have been realized without the generous support of Adrien Butty (Qualitas AG), Nicolas Berger (Swiss Herdbook), Martin Rust (Braunvieh Schweiz) and Timothée Neuenschwander (Holstein Schweiz) who granted access to the full database of day milk samples collected by Swiss dairy cow breeding associations over the past 40 years. In addition, I extend our appreciation to Peter Molnar from the ETH Chair of Hydrology and Water Resources Management for generously providing access to gridded weather data.

\vspace*{\baselineskip}
I would also like to thank Paul Rosenfield from Farmers Business Network who instructed and unblocked us with his in-depth knowledge in Generalized Additive Mixed Models with respect to the particular structure of our data. Without him we would probably not have been able to retrieve any meaningful results in such a short time horizon. In addition, Lukas Graz from the Consulting Service at the ETH Seminar of Statistics also provided valuable feedback for our questions. On top of that, I wish to acknowledge the insightful conversation with Laurence Jungo from Suisselab AG, who provided valuable insights regarding the history and methods of milk sampling in Switzerland.

\vspace*{\baselineskip}
Last but not least, I would like to thank Emma Lindberg, Brigitte Dorn, Maria Rey, and Emmanuel Frossard - the study administration of the Institute of Agricultural Sciences at ETH - for their support over the past years.